---
title: "MagicTCG"
output: html_notebook
Author: Nicko Saveas
---

# Setup

```{r setup}
library("rjson")
setwd("C:/Users/Nicko/Documents/School/Spring_2018/CS_422_Data_Mining/Data_Mining/MagicTCG/")

all_sets <- fromJSON(file="./AllSets.json")
```

all_sets
  set_name - each set is at this level i.e. "AKH"
    name
    code
    releaseData
    block
    border
    type
    booster
      how the booster is strucutred
      rare
      uncommmon
      etc
    mkm_name
    mkm_id
    cards[[1]] - each card of the set is at this level
    cards[[2]]
      artist
      cmc
      colorIdentity
      colors
      id
      name
      layout
      legalities
      manacost
      multiverseid
      name
      number
      power
      rarity
      subtext
      types
      toughness
      type
      types


# Dissecting Amonkhet

This is mostly just a test run so I can get the data in a format that is workable

```{r cleaning_trimming_and_transforming}
# These are really the only thing about that card that I care about
# The omitted attributes deal with legality and other things
card_attributes <- c("name", "type", "types", "colors", "convertedManaCost",  
                     "manaCost", "power", "toughness", "rarity", "originalText")

# Create a dataframe that contains all the cards
# Each row represents a card and each column and attribute of that card
amonkhet_cards.df <- setNames(data.frame(matrix(ncol = length(card_attributes), nrow = 0)), card_attributes)

# Iterate over all the cards in the set(s)
#for (i in 1:length(all_sets$AKH$cards))
#{
  # Create an empty row that we can fill
  amonkhet_cards.df[nrow(amonkhet_cards.df) + 1,] <- NA

  # Grab the card that we want given the index so we can iterate over it
  current_card <- all_sets$AKH$cards[[1]][card_attributes]
  
  # use the !is.null() fuction or the is.null() version

  # Now iterate over the card attributes and put them their values
  # into the row to populate one card (one row)
  for (j in 1:length(card_attributes))
  {
    # populate the row with each of the elements
     amonkhet_cards.df[nrow(amonkhet_cards.df), j] <- current_card[j]
  }
  
#}
```

```{r playing with data}



amonkhet_card_2 <- amonkhet$cards[[2]][card_attributes]
```




























