---
title: "MagicTCG"
output: html_notebook
Author: Nicko Saveas
---

# Setup

```{r setup}
library("rjson")
setwd("C:/Users/Nicko/Documents/School/Spring_2018/CS_422_Data_Mining/Data_Mining/MagicTCG/")

all_sets <- fromJSON(file="./AllSets.json")
```

all_sets
  set_name - each set is at this level i.e. "AKH"
    name
    code
    releaseData
    block
    border
    type
    booster
      how the booster is strucutred
      rare
      uncommmon
      etc
    mkm_name
    mkm_id
    cards[[1]] - each card of the set is at this level
    cards[[2]]
      artist
      cmc
      colorIdentity
      colors
      id
      name
      layout
      legalities
      manacost
      multiverseid
      name
      number
      power
      rarity
      subtext
      types
      toughness
      type
      types


# Dissecting Amonkhet

This is mostly just a test run so I can get the data in a format that is workable
```{r dissecting_amonkhet}
amonkhet <- all_sets$AKH
amonkhet_cards <- amonkhet$cards
amonkhet_card_1 <- amonkhet_cards[[1]]
```

```{r god_help_me}
amonkhet_card_1$name
amonkhet_card_1$type
amonkhet_card_1$colors
amonkhet_card_1$convertedManaCost
amonkhet_card_1$power
amonkhet_card_1$toughness
amonkhet_card_1$rarity
```

```{r playing with data}

```




























