---
title: "CS 422 Section 02"
output: html_notebook
author: Nicholas Saveas
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 1.1

I am taking this course because I feel that a career in data science would allow me to make use of both my computer science and math skills. I would like to achieve a greater understanding of what the field of data science is about and am using this course to help me determine what I would like to do with my future. 

# Exercise 2.1

## 2.1-A

Load the dataset using "data(cars)" and then print it by calling it.

```{r}
data(cars)
cars
```

## 2.1-B

Plot the cars dataset using the "plot(cars)" comand and then provide the graph with a "main = _" title.

```{r}
plot(cars)
title(main = 'The Cars Dataset')
```

## Exercise 2.1-C

The summary() function gives you enough information to create a box plot - minimum, 1st quartile, median, 3rd quartile, and maximum. + the mean

```{r}
summary(cars)
```

## Exercise 2.1-D

The maximum speed is 25.0  

The minimum distance is 2.00

# Exercise 2.2

## Exercise 2.2-A

Set the working directory to where the "students.csv"" is stored.
Load the dataframe and then add a new element to the dataset. 

stringsAsFactors = TRUE treats strings as enumerables - categorical variable. It has to be a value in a certain limited number of variables.


```{r}
setwd("C:/Users/Nicko/Documents/School/Spring_2018/CS_422_Data_Mining/Data_Mining/Basic_R")
dataFrame <- read.csv("student.csv", sep=";", header=T, stringsAsFactors = FALSE)
dataFrame <- rbind(dataFrame, c("Brad Pitt", "40051", "2.21", "C"))
dataFrame
```
